### 2021-03-22 README.Rmd
Update README to explain all that it does so far.

---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "80%",
  fig.align = "center"
)
```

# Steve's Academic R Projects

```{r echo=FALSE, results="hide", message=FALSE}
library("badger")
```

```{r, echo = FALSE, results='asis'}
cat(
	badge_cran_release("steveproj", "green"),
	badge_cran_download("steveproj", "grand-total", "green"),
	badge_cran_download("steveproj", "last-month", "green"),
	badge_cran_download("steveproj", "last-week", "green")
)
```

<img src="http://svmiller.com/images/steveproj-hexlogo.png" alt="My steveproj hexlogo" align="right" width="200" style="padding: 0 15px; float: right;"/>

`{steveproj}` is an R package to help you start and manage R projects with an eye toward producing an academic project (to culminate in an academic paper). The package is in development and will incorporate other packages in my eponymous R ecosystem, prominently [`{stevetemplates}`](https://github.com/svmiller/stevetemplates). The package itself leans primarily on Rstudio and how it creates/manages projects. The creation a project in Rstudio, through `{steveproj}`, will result in a new directory containing subdirectories for rendering documents (`src`), R scripts for analysis (`R`), finished data objects (`data`), and finished reports (`doc`). An accompanying `Makefile` and R Markdown file will assist in the management and production of the project.

## What It Does (So Far)

Assuming the latest/development version of [`{stevetemplates}`](https://github.com/svmiller/stevetemplates) (i.e. the one that has the Word template) and considering any potential LaTeX weirdness that comes from different builds, this package has function in R and Make that will:

- simplify the research process to three basic "targets" (in Make): a finished data product to analyze, statistical models of the data, and post-estimation simulations of quantities of interest from the data. These scripts are in the `R/` directory and render to the `data/` directory.
- render your R Markdown document to a fancy PDF document in LaTeX using [my second article template](http://svmiller.com/blog/2020/09/another-rmarkdown-article-template/).
- render your R Markdown document to an anonymized version of that same document. [A post on my blog](http://svmiller.com/blog/2021/03/handle-academic-projects-steveproj-make/) gives clues how to do this with YAML parameters.
- render your R Markdown document to an anonymized Word document. Anonymized Word documents are all that I'm willing to support here. The goal isn't to publish to Word, per se; it's only to produce a document suitable for peer review for journals that demand you provide one.
- farm your R Markdown document for citations and format them to a bibliography file. By default, this will render to a `refs.bib` file in the `inst/` directory. You can tweak this if you like.
- scan your R Markdown document to render a simple title page for peer review. Traditionally, journals ask for a title page (with author information) and a manuscript (without author information). `render_abstract.R` in the `src/` directory is the companion script to `render_pdf-anon.R` in the same directory.

## Installation

When the time comes, you can install this on CRAN.

```r
install.packages("steveproj")
```

Right now, this package in development and is not available on CRAN. You can install the development version of `steveproj` from Github via the `devtools` package. I suppose using the `remotes` package would work as well.

```r
devtools::install_github("svmiller/steveproj")
```

## Usage

My website will have a more exhaustive tutorial for using this package and all that comes in it. For now, a user maximize their experience with `{steveproj}` through either the console or Rstudio.

### Console

Assume the current working directory is something like `/home/steve/Dropbox/projects` (as it is for me). Then, a user can create a new project with the `create_project()` command.

```{r, eval=F}
steveproj::create_project("dissertation")
```

This will create a new directory in the working directory, titled "dissertation", that includes a skeleton of a research project to assist the user in getting started on their academic paper/project. That directory will be located in the current working directory. The only downside to the console approach over the Rstudio approach is the console approach won't create an `.Rproj` file in the directory. The Rstudio approach will.

You could optionally combine this command line call in the console with the following command, which will add an `.Rproj` file to the directory you just created

```{r, eval=F}
# steveproj::create_project("dissertation")
# ^ assume you just ran this
# Then, do this next
steveproj::create_rproj("dissertation")
```

### Rstudio

Go to *File > New Project*. You'll see a prompt that looks like this. Select "New Directory".

![](inst/README/steveproj-1.png)

That will direct you here. Scroll down your available project types until you see my "S" icon, which is incidentally the favicon on [my website](http://svmiller.com). Select that entry to create a new academic paper/project.

![](inst/README/steveproj-2.png)

That will direct you here. Enter the name of the directory you want to create. Click "Create Project" when you're done. This will create a new directory, titled whatever you entered in the directory name, along with an `.Rproj` file. By default, it will also open a new Rstudio session.

![](inst/README/steveproj-3.png)