### 2021-11-18 inst/other_docs/memo.Rmd

---
geometry: margin=1in
header-includes:
  - \linespread{1.05}

author: "&nbsp;"
address: ""
email: ""
url: ""
phone: ""
fontsize: 11pt
# mainfont: cochineal
# monofont: Fira Code

date: "`r format(Sys.time(), '%e %B %Y')`"
greetings: "Dear Editor and Reviewers,"
# bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`"
# biblio-style: apsr
---

```{r setup, include=F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(tidyverse)
library(peacesciencer)
library(kableExtra)

is_docx <- knitr::pandoc_to("docx")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

```

I thank you all for the opportunity to submit revisions of my manuscript "[INSERT YOUR MANUSCRIPT NAME HERE]" for possible inclusion in *INSERT JOURNAL HERE*. I thank the [two/three/four/five/six/seven/eight] reviewers and the editorial team for the helpful comments as well. In what follows, I address the editor about the changes to the manuscript that came in light of reviewer feedback. Thereafter, I address the reviewers' comments individually.

# Comments to the Editorial Team

Do you ever feel like a plastic bag? Drifting through the wind, wanting to start again?

Best regards,

The Author

\newpage

# Reviewer #1's Comments

Reviewer #1 (R1) said some things. I said some things. We all agreed to move past it.

# Reviewer #2's Comments

I thank Reviewer #2 (R2) for their comments. R2 felt I was [2 legit 2 quit](https://www.youtube.com/watch?v=HFCv86Olk8E). That was a very sweet thing for R2 to say and I thank them for it.